---
title: "nest_spacing"
author: "Chris Hill"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in data into a file called "data" and show its structure
- before you can execute this you have to save the nest card data as a csv and name it "nest_card_dataYYYYMMDD.csv" The filename has to match what's in the code below.
- Then put the csv vile in Users/chill/github/nest_spacing/ (or your own appropriate directory if you're not me)
- since nest card data is a big sprawling data file the resulting object in r is kind of a mess, see below.

```{r read}
data<-read.csv("nest_card_data20220728.csv")
str(data)
```

## Rename columns
- just renaming NEST_TERRITORY because it's too long.
```{r}
library(dplyr)
data <- dplyr::rename(data, TERRITORY = NESTING_TERRITORY)
```


## Filter data
- keep only the valid nests and put them in "data2"
- confirm that only nests with "TRUE" are left (*yes, it worked*)
- I could also subset to just certain years at this point
```{r}
  data2<-subset(data, VALID_NEST==TRUE) #remove all but type 2 visits
unique(data2$VALID_NEST) #what's left
```

## Get rid of columns I don't need
- keep 7 columns in an object called "data3"
- show what's left using "head" and "str"
- Hatch date is not recognized as dates here, it's "chr" data still.
```{r subset}
data3 <- subset(data2, select = c(NEST_NUMBER,TERRITORY,VALID_NEST,LAT,LONG,HATCH_DATE, HD_YEAR))

head(data3)#what's left
```
```{r}
str(data3)
```

## Convert date from character class to r date format
- then check with "sapply"


```{r}
library(lubridate)
data3$HATCH_DATE <- lubridate::dmy(data3$HATCH_DATE) #convert date
sapply(data3, class) #check classes to make sure date was reformatted
```



