---
title: "nest_spacing"
author: "Chris Hill"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in data into a file called "data" and show its structure
- before you can execute this you have to save the nest card data as a csv and name it "nest_card_dataYYYYMMDD.csv" The filename has to match what's in the code below.
- Then put the csv vile in Users/chill/github/nest_spacing/ (or your own appropriate directory if you're not me)
- since nest card data is a big sprawling data file the resulting object in r is kind of a mess, see below.

```{r read}
data<-read.csv("nest_card_data20220729.csv")
str(data)
```

## Rename columns
- just renaming NEST_TERRITORY because it's annoyingly long.
```{r}
library(dplyr)
data <- dplyr::rename(data, TERRITORY = NESTING_TERRITORY)
```


## Filter data into "data2"
- keep only the valid nests and put them in "data2"
- confirm that only nests with "TRUE" are left (*yes, it worked*)
- I could also subset to just certain years at this point
```{r}
  data2<-subset(data, VALID_NEST==TRUE) #remove all but type 2 visits
unique(data2$VALID_NEST) #what's left
```

## Get rid of columns I don't need to make "data3"
- keep 7 columns in an object called "data3"
- show what's left using "head" and "str"
- Hatch date is not recognized as dates here, it's "chr" data still.
```{r subset}
data3 <- subset(data2, select = c(NEST_NUMBER,TERRITORY,VALID_NEST,LAT,LONG,HATCH_DATE, HD_YEAR))

head(data3)#what's left
```
```{r}
str(data3)
```

## Convert date from character class to r date format
- then check with "sapply"
- one date is still not parsing (I though getting rid of invalid nests would solve that, but I guess not)
- write data3 to a file on my hard drive


```{r}
library(lubridate)
data3$HATCH_DATE <- lubridate::dmy(data3$HATCH_DATE) #convert date
sapply(data3, class) #check classes to make sure date was reformatted
write.csv(data3, "data3.csv", row.names = FALSE)
```


## To calculate distances between points I am going to install the package "sp" and use the function spDistsN1().
- description: The function returns a vector of distances between a matrix of 2D points, first column longitude, second column latitude, and a single 2D point, using Euclidean or Great Circle distance (WGS84 ellipsoid) methods.
- wait maybe there's a better way, see below for a more complete solution (I'm commenting out the "sp" stuff for now)


```{r}
#install.packages("sp")
#library("sp")
```
## Another source suggests using packages sp and rgeos so I will load rgeos, too. Or maybe sf is the best way.  I'll try sf first.
- link to post: https://stackoverflow.com/questions/21977720/r-finding-closest-neighboring-point-and-number-of-neighbors-within-a-given-rad
- The documentation on rgeos says "Please note that rgeos will be retired by the end of 2023,
plan transition to sf functions using GEOS at your earliest convenience." So I will load sf.  I guess I'll try to implement the solution using sf even though I don't understand a lot of the basics of the reprex.  For now the "install.packages("sf") is commented out because I was already playing with it.
```{r}
#install.packages("rgeos")
#library(rgeos)
#install.packages("sf")
library(sf)
```
## Load the data in data3 into data4, a "sf object" 
- "sf", for "simple features", is a format for containing both geographic information and attributes in the same file. It's a special data frame where geographic information is in a special column, and datum is specified as well.


```{r}
data4 <- st_as_sf(data3, coords = c("LAT", "LONG"), 
crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
```
